---
title: "TextDescribR Development"
author: "Jamie Reilly, Shawn Wang, Avery Pattullo"
date: "`r format(Sys.Date(), '%B %Y')`"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    toc_depth: 1
    css: "https://github.com/Reilly-ConceptsCognitionLab/reillylab_publicdata/blob/main/jr_20.css?raw=true"
---

```{r setup, include=F}
knitr::opts_chunk$set(fig.width=8, fig.height=5, fig.align = 'left', fig.path='Figs/', cache.path='Cache/', eval=T, echo=T, tidy=TRUE,  cache=F, message=F, warning=F, stringsAsFactors=F, yaml.eval.expr = TRUE)  

library(xfun)
pkg_attach("tidyverse", "here", "kableExtra", "RCurl", "psych", "knitr", "RColorBrewer", 'dplyr', 'tidyr', install=T)

jamie.theme <- theme_bw() + theme(axis.line = element_line(colour = "black"), panel.grid.minor = element_blank(), panel.grid.major = element_blank(), panel.border = element_blank(), panel.background = element_blank(), legend.title= element_blank())  #custom theme ggplot2


print.me <- function(x, ...) {
if (nrow(x) > 200){
   len <- 200 
    } else {
       len <- (nrow(x))
}
   x[1:len,] %>%
   kbl(digits=2, align= 'l', booktabs=T) %>%
   kable_styling(fixed_thead = T) %>%
   kable_paper("striped", full_width = T, html_font = "Helvetica", font_size = 12) %>%
   row_spec(0, color = "white", background = "#5b705f", font_size = 12) %>%
   scroll_box(width = "700px", height = "500px") %>%
   asis_output()
}

registerS3method("knit_print", "data.frame", print.me)
#to_R <- read.csv(here("data", "MyRaw.txt")) ---- using 'here' to read in data
```

# Load test texts
```{r}
#Test_Text1 <- read.csv('data/Test_Text1.csv')
#Save and post to github reillylab_public
#save(Test_Text1, file='data/Test_Text1.rda')

#Load Test_Text1
load(url("https://github.com/Reilly-ConceptsCognitionLab/reillylab_publicdata/blob/main/Test_Text1.rda?raw=true"))
str(Test_Text1)


#Load Test_Text2

```

# Variables for Summary Table
1. Document count <br>
2. Total word/token count (all documents) <br>
2. Word count per-document (mean, sd, range) <br>
3. Sentence count per-document (mean, sd, range) <br>
4. Sentence length (in words) (mean, sd, range) <br>
5. Stopword proportion (mean, sd, range) <br>
6. TTR (mean, sd, range) <br>
7. word frequency (lg10) (mean, sd, range, %missing vals)
8. word concreteness (mean, sd, range, %missing vals)
9. word valence  (mean, sd, range, %missing vals)
10. word complexity by Flesch-Kincaid (mean, sd, range, %missing vals)

#Logic
We will create a large, long form dataframe the original text. This datafranme will contain
all stats listed above as columns and identifiers for document, sentence and word in case someone wants stats at the document level instead of corpus level.


# Define Cleaning and Computation functions
##Load replacements_25
In the real package this function will be internal and loaded automatically in the background.
```{r}
replacements_25 <- function(dat, wordcol) {
  # Apply all replacements in sequence
  dat %>%
    # Contractions starting with a/i
    dplyr::mutate({{wordcol}} := gsub("\\baren't\\b", "are not", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bcan't\\b", "cannot", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bcould've\\b", "could have", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bcouldn't\\b", "could not", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bdidn't\\b", "did not", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bdoesn't\\b", "does not", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bdon't\\b", "do not", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bdunno\\b", "do not know", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bgimme\\b", "give me", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bgonna\\b", "going to", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bgotta\\b", "got to", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bhadn't\\b", "had not", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bhasn't\\b", "has not", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bhaven't\\b", "have not", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bhe'll\\b", "he will", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bhe's\\b", "he is", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bi'd\\b", "i would", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bi'll\\b", "i will", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bi'm\\b", "i am", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bi've\\b", "i have", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bisn't\\b", "is not", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bit'll\\b", "it will", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bit's\\b", "it is", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bkinda\\b", "kind of", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\blemme\\b", "let me", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bmight've\\b", "might have", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bmightn't\\b", "might not", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bmust've\\b", "must have", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bmustn't\\b", "must not", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bneedn't\\b", "need not", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bshe'll\\b", "she will", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bshe's\\b", "she is", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bshould've\\b", "should have", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bshouldn't\\b", "should not", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bsorta\\b", "sort of", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bthat's\\b", "that is", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bthere'd\\b", "there would", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bthey'll\\b", "they will", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bthey're\\b", "they are", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bthey've\\b", "they have", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bwanna\\b", "want to", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bwasn't\\b", "was not", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bwe'll\\b", "we will", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bwe're\\b", "we are", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bwe've\\b", "we have", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bweren't\\b", "were not", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bwon't\\b", "will not", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bwould've\\b", "would have", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bwouldn't\\b", "would not", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\byou'd\\b", "you would", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\byou'll\\b", "you will", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\byou're\\b", "you are", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\byou've\\b", "you have", {{wordcol}}, ignore.case = TRUE))
}

```

## Clean_1 (Sentence Level)
Cleans and tokenizes by sentence to derive sentence level analytics. Formats dataframe after doing initial cleaning. Generates the following stats: <br>
1) Total documents <br>
2) Sentences per document (mean, sd, range) <br>
3) Words per sentence per document (mean, sd, range) <br>
4) Words per document (mean, sd, range) <br>

Considerations and questions to resolve: <br>
1) Output two dataframes as a list, One df is the raw corpus split in tidyformat one word per row with several columns of metadata (doc_id, sentence_num, word_num). The other is an empty data table that will get populated with corpus analytics then converted into a journal quality table <br>

To do steps (not executed yet): <br>
1) Add hyphen and em dash replacements
1) Check to make sure that word count per sentence is correct <br>
2) Make sure word sequence and sent sequence are correct <br>
3) make sure replacements are working <br>


```{r}
clean_1 <- function(dat, text_field='my_text', doc_id_field='doc_id') {
  #error message to user if failed to specify doc_ID_field
  if (is.null(doc_id_field) || !doc_id_field %in% colnames(dat)) {
    stop("Oops... it's important that you tell TextDescribR what column corresponds to your document_ID as an argument. Remember to enclose your column header in quotes like doc_id_field='my_docs'")
  }
  
  #error message to user if failed to specify text_field
  if (is.null(text_field) || !text_field %in% colnames(dat)) {
    stop("Oops... it's important that you tell TextDescribR what column corresponds to your text_field as an argument. Remember to enclose your column header in quotes like text_field='my_text'")
  }
  
  # Select and prepare data
  x <- dat
  
  #standrdize text field contents to UTF-8 encoding first
  x[[text_field]] <- iconv(x[[text_field]], from = "", to = "UTF-8")

# Convert doc_ID_field to factor
  x[[doc_id_field]] <- as.factor(x[[doc_id_field]])

  # convert text_id field to character
x[[text_field]] <- as.character(x[[text_field]])

# standarize apostrophes
# text to to lowercase
# Sub common symbols for full words (space before each word, e.g. % for '_percent', 
# Change elipses to nothing (2 or more adjacent periods with closed space)
# Sub hyphens for space
 x <- x %>% mutate(!!sym(text_field) := stringi::stri_replace_all_regex(!!sym(text_field), "[\u2018\u2019\u02BC\u201B\uFF07\u0092\u0091\u0060\u00B4\u2032\u2035]", "'")) %>%
  mutate(!!sym(text_field) := tolower(!!sym(text_field))) %>%
  mutate(!!sym(text_field) := str_replace_all(!!sym(text_field), "%", " percent")) %>%
  mutate(!!sym(text_field) := str_replace_all(!!sym(text_field), "\\+", " plus")) %>%
  mutate(!!sym(text_field) := str_replace_all(!!sym(text_field), "\\&", " and")) %>%
  mutate(!!sym(text_field) := str_replace_all(!!sym(text_field), "\\.{2,}", "")) %>%
  mutate(!!sym(text_field) := str_replace_all(!!sym(text_field), "\\-", " ")) 
 
 #Expand contractions
#Include function as internal 'repolacements_25` in package
# Apply contractions replacement do NOT quote the column name
x <- replacements_25(dat = x, wordcol = !!sym(text_field))
  
#Remove decimals via regex, PERL allows lookaheads/lookbehinds - (?<=\\d) means lookbehind for one or more digits
# Replaces 2 or more adjacent spaces with one space
x <- x %>% mutate(!!sym(text_field) := sub("(?<=\\d)\\.(?=\\d)", "", !!sym(text_field), perl = TRUE)) %>% mutate(!!sym(text_field) := gsub("\\s{2,}", " ", !!sym(text_field)))
 
 #tokenize (split) sentences one per row using delimiters (?, ., !)
x <- x %>% dplyr::mutate(!!sym(text_field) := stringr::str_split(!!sym(text_field), "[.!?]", simplify = FALSE)) %>% tidyr::unnest(!!sym(text_field))

#Omit empty rows, Not NA & Not an Empty String
x <- x %>% dplyr::filter(!is.na(!!sym(text_field)) & trimws(!!sym(text_field)) != "")

#Add `sent_seq` sequence variable looping over each document (starts new at switch, break on factor level change for doc_id)
x <- x %>% dplyr::group_by(!!sym(doc_id_field)) %>% dplyr::mutate(ID_SentSeq_ByDoc = row_number()) %>% dplyr::ungroup()

#sent_seq as factor (count levels of factor by document for sentence counts)
x$ID_SentSeq_ByDoc <- as.factor(x$ID_SentSeq_ByDoc)

#Omit all non-alphabetic characters
x <- x %>% mutate(!!sym(text_field) := str_replace_all(!!sym(text_field), "[[:punct:]]", "")) 

#Generate word count per sentence, any non-whitespace character is counted as a word
#MAKE SURE THIS WORKS
x <- x %>% mutate(Stat_WordCount_BySent = sapply(!!sym(text_field), function(text) str_count(text, "\\S+")))    

#group by doc_id and SentSeqID_ByDoc, counts the number of distinct levels of SentSeqID_ByDoc output as total sentences per doc
x <- x %>% dplyr::group_by(!!sym(doc_id_field)) %>% dplyr::mutate(Stat_SentCount_ByDoc = dplyr::n_distinct(ID_SentSeq_ByDoc)) %>% dplyr::ungroup()

#Split and unlist data to one word-per row format 
x <- x %>% tidyr::separate_rows(!!sym(text_field), sep = " ")

#Omit empty rows AGAIN (missing and empty strings)
x <- x %>% dplyr::filter(!is.na(!!sym(text_field)) & !!sym(text_field) != "")

# Generate Word_Num variable looping down dataframe, break and restart on new factor level of doc_id
x <- x %>% dplyr::group_by(!!sym(doc_id_field)) %>% dplyr::mutate(ID_WordSeq_ByDoc = row_number()) %>% dplyr::ungroup()
 
#Generate word-count-per sentence (total rows in each level of sentence)
x <- x %>% group_by(!!sym(doc_id_field), ID_SentSeq_ByDoc) %>% mutate(Stat_WordCount_BySent = n()) %>% ungroup()

# Generate total words per document (total rows undifferentiated by sentence number)
x <- x %>% dplyr::group_by(!!sym(doc_id_field)) %>% dplyr::mutate(Stat_WordCount_ByDoc = n()) %>% 
dplyr::ungroup()
 
#Count number of documents (levels of factor doc_id_field)
x$Stat_DocCount <- nlevels(x[[doc_id_field]])

#reorder dataframe so ID vars come before stats, move doc_id second
#x <- x[2, 1, "ID_SentSeq_ByDoc", "ID_WordSeq_ByDoc"]

#bind stats and dataframe into list
#clean1_list <- ls(stat1_doc_count)
 
#return(clean1_list)
return(x)
}
```


# Test clean_1
```{r}
try1 <- clean_1(Test_Text1, text_field='my_text', doc_id_field='doc_id')
head(try1, n=75)
colnames(try1)
#write.csv(try1, "output/try_it_out.csv", row.names=F)
```

## clean_2
computes all other lexical level analytics on each word (one-per-row). <br>

**do**:
1. Load lookup database from ConversationAlign (don't store data in package) <br>
```{r}

```

#
#
#
#
# RESOLVED
1. Add informative error messages explaining to user `doc_id` and `my_text` fields <br>
2. Generate `sentence_num` using seq_along in the first steps after tokenizing by sentence but before splitting into one word per row format <br>
3. Generate `word_num` using seq_along after splitting and unlisting. Make sure to generate this on the grouped data (word_num sequentially ordered by document) <br>
4. Count number of documents in corpus using levels, write to empty dataframe <br>
5. Read data, add error message if user does not specify correct text fields and doc_ids <br>
6. Convert `doc_id_field` to factor, group by `doc_id_field` for summary stats <br>

