---
title: "TextDescribR Development"
author: "Jamie Reilly, Shawn Wang, Avery Pattullo"
date: "`r format(Sys.Date(), '%B %Y')`"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    toc_depth: 1
    css: "https://github.com/Reilly-ConceptsCognitionLab/reillylab_publicdata/blob/main/jr_20.css?raw=true"
---

```{r setup, include=F}
knitr::opts_chunk$set(fig.width=8, fig.height=5, fig.align = 'left', fig.path='Figs/', cache.path='Cache/', eval=T, echo=T, tidy=TRUE,  cache=F, message=F, warning=F, stringsAsFactors=F, yaml.eval.expr = TRUE)  

library(xfun)
pkg_attach("tidyverse", "here", "kableExtra", "RCurl", "psych", "knitr", "RColorBrewer", 'dplyr', install=T)

jamie.theme <- theme_bw() + theme(axis.line = element_line(colour = "black"), panel.grid.minor = element_blank(), panel.grid.major = element_blank(), panel.border = element_blank(), panel.background = element_blank(), legend.title= element_blank())  #custom theme ggplot2


print.me <- function(x, ...) {
if (nrow(x) > 200){
   len <- 200 
    } else {
       len <- (nrow(x))
}
   x[1:len,] %>%
   kbl(digits=2, align= 'l', booktabs=T) %>%
   kable_styling(fixed_thead = T) %>%
   kable_paper("striped", full_width = T, html_font = "Helvetica", font_size = 12) %>%
   row_spec(0, color = "white", background = "#5b705f", font_size = 12) %>%
   scroll_box(width = "700px", height = "500px") %>%
   asis_output()
}

registerS3method("knit_print", "data.frame", print.me)
#to_R <- read.csv(here("data", "MyRaw.txt")) ---- using 'here' to read in data
```

# Load test texts
```{r}
#Test_Text1 <- read.csv('data/Test_Text1.csv')
#Save and post to github reillylab_public
#save(Test_Text1, file='data/Test_Text1.rda')

#Load Test_Text1
load(url("https://github.com/Reilly-ConceptsCognitionLab/reillylab_publicdata/blob/main/Test_Text1.rda?raw=true"))
str(Test_Text1)
```

# Possible variables
1. Document count
2. Total word count per document (mean, sd, range)
3. Total sentence count per document (mean, sd, range)
4. Words-per-sentence per document (mean, sd, range)
5. % Stopwords per document (mean, sd, range)
6. TTR per document


# Clean_1
Function does some cleaning, tokenizing by sentence to derive sentence level analytics (sentence count, words per sentence). IMPORTANT -- keep periods but not when the periods are decimal points.
<br>

Considerations and questions to resolve: <br>
1) Read corpus, if multidocument group_by `doc_id` <br>
2) Add informative error messages explaining to user `doc_id` and `my_text` fields <br>
3) Output two dataframes as a list, One df is the raw corpus split in tidyformat one word per row with several columns of metadata (doc_id, sentence_num, word_num). 
4) Generate `sentence_num` using seq_along in the first steps after tokenizing by sentence but before splitting into one word per row format <br>
5) Generate `word_num` using seq_along after splitting and unlisting. Make sure to generate this on the grouped data (word_num sequentially ordered by document) <br>

Planned steps:
1) Read data, add error message if user does not specify correct text fields and doc_ids
2) Convert `doc_id` to factor
3) Count number of documents in corpus using levels, write to empty dataframe 
4) Group by `doc_id_field` whatever the user calls it


```{r}

clean_1 <- function(dat, text_field='my_text', doc_id_field='doc_id') {
  #error message to user if failed to specify doc_ID_field
  if (is.null(doc_id_field) || !doc_id_field %in% colnames(dat)) {
    stop("Oops... it's important that you tell TextDescribR what column corresponds to your document_ID as an argument. Remember to enclose your column header in quotes like doc_id_field='my_docs'")
  }
  
  #error message to user if failed to specify text_field
  if (is.null(text_field) || !text_field %in% colnames(dat)) {
    stop("Oops... it's important that you tell TextDescribR what column corresponds to your text_field as an argument. Remember to enclose your column header in quotes like text_field='my_text'")
  }
  
  #standrdize text field contents to UTF-8 encoding first
  dat[[text_field]] <- iconv(dat[[text_field]], from = "", to = "UTF-8")
  
  # Select and prepare data
  x <- dat
  
  # Convert the specified document ID field to factor
  x[[doc_id_field]] <- as.factor(x[[doc_id_field]])
  
  #grouping by doc_id_field
  #TBD
  
  # Convert the text to lowercase
  x[[text_field]] <- tolower(x[[text_field]])
  
  #Remove decimals
  #Remove decimals via regex, PERL allows lookaheads/lookbehinds - 
  # (?<=\\d) means lookbehind for one or more digits
 x <- x %>% mutate(!!sym(text_field) := sub("(?<=\\d)\\.(?=\\d)", "", !!sym(text_field), perl = TRUE))

 #tokenize (split) sentences one per row using delimiters (?, ., !)
 
 
 #Drop sent_num sequence variable looping over each document (starts new at switch)
 
 #write results to empty dataframe which we will continue to populate dataframe_1
 
 #split and unlist data and write to dataframe_2
 
 #bind dataframe_1 and dataframe_2 into a list
 
 #return the list of two dataframes

  return(x)
}
```


# Test clean_1
```{r}
try_it_out <- clean_1(Test_Text1, text_field='my_text', doc_id_field='doc_id')
print(try_it_out)
```




