---
title: "TextDescribR Development"
author: "Jamie Reilly, Shawn Wang, Avery Pattullo"
date: "`r format(Sys.Date(), '%B %Y')`"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    toc_depth: 1
    css: "https://github.com/Reilly-ConceptsCognitionLab/reillylab_publicdata/blob/main/jr_20.css?raw=true"
---

```{r setup, include=F}
knitr::opts_chunk$set(fig.width=8, fig.height=5, fig.align = 'left', fig.path='Figs/', cache.path='Cache/', eval=T, echo=T, tidy=TRUE,  cache=F, message=F, warning=F, stringsAsFactors=F, yaml.eval.expr = TRUE)  

library(xfun)
pkg_attach("tidyverse", "here", "kableExtra", "RCurl", "psych", "knitr", "RColorBrewer", 'dplyr', install=T)

jamie.theme <- theme_bw() + theme(axis.line = element_line(colour = "black"), panel.grid.minor = element_blank(), panel.grid.major = element_blank(), panel.border = element_blank(), panel.background = element_blank(), legend.title= element_blank())  #custom theme ggplot2


print.me <- function(x, ...) {
if (nrow(x) > 200){
   len <- 200 
    } else {
       len <- (nrow(x))
}
   x[1:len,] %>%
   kbl(digits=2, align= 'l', booktabs=T) %>%
   kable_styling(fixed_thead = T) %>%
   kable_paper("striped", full_width = T, html_font = "Helvetica", font_size = 12) %>%
   row_spec(0, color = "white", background = "#5b705f", font_size = 12) %>%
   scroll_box(width = "700px", height = "500px") %>%
   asis_output()
}

registerS3method("knit_print", "data.frame", print.me)
#to_R <- read.csv(here("data", "MyRaw.txt")) ---- using 'here' to read in data
```

# Load test texts
```{r}
#Test_Text1 <- read.csv('data/Test_Text1.csv')
#Save and post to github reillylab_public
#save(Test_Text1, file='data/Test_Text1.rda')

#Load Test_Text1
load(url("https://github.com/Reilly-ConceptsCognitionLab/reillylab_publicdata/blob/main/Test_Text1.rda?raw=true"))
str(Test_Text1)
```

# Possible variables
1. Document count <br>
2. Word count per-document (mean, sd, range) <br>
3. Sentence count per-document (mean, sd, range) <br>
4. Word count per-sentence per-document (mean, sd, range) <br>
5. Proportion of stopwords per-document (mean, sd, range) <br>
6. TTR per-document <br>


# Define Cleaning and Computation functions
## Clean_1 (Sentence Level)
Cleans and tokenizes by sentence to derive sentence level analytics (sentence count, words per sentence). <br> 

Considerations and questions to resolve: <br>
1) Read corpus, if multidocument group_by `doc_id` <br>
2) Output two dataframes as a list, One df is the raw corpus split in tidyformat one word per row with several columns of metadata (doc_id, sentence_num, word_num). 

Planned steps:
1) Read data, add error message if user does not specify correct text fields and doc_ids
2) Convert `doc_id_field` to factor, group by `doc_id_field` for summary stats

#Load replacements_25
In the real package this function will be internal and loaded automatically in the background.
```{r}
replacements_25 <- function(dat, wordcol) {
  # Apply all replacements in sequence
  dat %>%
    # Contractions starting with a/i
    dplyr::mutate({{wordcol}} := gsub("\\baren't\\b", "are not", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bcan't\\b", "cannot", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bcould've\\b", "could have", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bcouldn't\\b", "could not", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bdidn't\\b", "did not", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bdoesn't\\b", "does not", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bdon't\\b", "do not", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bdunno\\b", "do not know", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bgimme\\b", "give me", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bgonna\\b", "going to", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bgotta\\b", "got to", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bhadn't\\b", "had not", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bhasn't\\b", "has not", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bhaven't\\b", "have not", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bhe'll\\b", "he will", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bhe's\\b", "he is", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bi'd\\b", "i would", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bi'll\\b", "i will", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bi'm\\b", "i am", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bi've\\b", "i have", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bisn't\\b", "is not", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bit'll\\b", "it will", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bit's\\b", "it is", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bkinda\\b", "kind of", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\blemme\\b", "let me", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bmight've\\b", "might have", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bmightn't\\b", "might not", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bmust've\\b", "must have", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bmustn't\\b", "must not", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bneedn't\\b", "need not", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bshe'll\\b", "she will", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bshe's\\b", "she is", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bshould've\\b", "should have", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bshouldn't\\b", "should not", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bsorta\\b", "sort of", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bthat's\\b", "that is", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bthere'd\\b", "there would", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bthey'll\\b", "they will", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bthey're\\b", "they are", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bthey've\\b", "they have", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bwanna\\b", "want to", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bwasn't\\b", "was not", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bwe'll\\b", "we will", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bwe're\\b", "we are", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bwe've\\b", "we have", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bweren't\\b", "were not", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bwon't\\b", "will not", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bwould've\\b", "would have", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\bwouldn't\\b", "would not", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\byou'd\\b", "you would", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\byou'll\\b", "you will", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\byou're\\b", "you are", {{wordcol}}, ignore.case = TRUE)) %>%
    dplyr::mutate({{wordcol}} := gsub("\\byou've\\b", "you have", {{wordcol}}, ignore.case = TRUE))
}

```





```{r}
clean_1 <- function(dat, text_field='my_text', doc_id_field='doc_id') {
  #error message to user if failed to specify doc_ID_field
  if (is.null(doc_id_field) || !doc_id_field %in% colnames(dat)) {
    stop("Oops... it's important that you tell TextDescribR what column corresponds to your document_ID as an argument. Remember to enclose your column header in quotes like doc_id_field='my_docs'")
  }
  
  #error message to user if failed to specify text_field
  if (is.null(text_field) || !text_field %in% colnames(dat)) {
    stop("Oops... it's important that you tell TextDescribR what column corresponds to your text_field as an argument. Remember to enclose your column header in quotes like text_field='my_text'")
  }
  
  #standrdize text field contents to UTF-8 encoding first
  dat[[text_field]] <- iconv(dat[[text_field]], from = "", to = "UTF-8")
  
# Select and prepare data
  x <- dat

# Convert the specified document ID field to factor
  x[[doc_id_field]] <- as.factor(x[[doc_id_field]])
  
# standarize apostrophes
# text to to lowercase
# Sub common symbols for full words (space before each word, e.g. % for '_percent'
# Change elipses to nothing (2 or more adjacent periods with closed space)
 x <- x %>% 
  mutate(!!sym(text_field) := stringi::stri_replace_all_regex(!!sym(text_field), "[\u2018\u2019\u02BC\u201B\uFF07\u0092\u0091\u0060\u00B4\u2032\u2035]", "'")) %>%
  mutate(!!sym(text_field) := tolower(!!sym(text_field))) %>%
  mutate(!!sym(text_field) := str_replace_all(!!sym(text_field), "%", " percent")) %>%
  mutate(!!sym(text_field) := str_replace_all(!!sym(text_field), "\\+", " plus")) %>%
  mutate(!!sym(text_field) := str_replace_all(!!sym(text_field), "\\&", " and")) %>%
  mutate(!!sym(text_field) := str_replace_all(!!sym(text_field), "\\.{2,}", ""))
  
#Remove decimals via regex, PERL allows lookaheads/lookbehinds - 
# (?<=\\d) means lookbehind for one or more digits
# Replaces 2 or more adjacent spaces with one space
 x <- x %>% mutate(!!sym(text_field) := sub("(?<=\\d)\\.(?=\\d)", "", !!sym(text_field), perl = TRUE)) %>% mutate(!!sym(text_field) := gsub("\\s{2,}", " ", !!sym(text_field)))
 
 #tokenize (split) sentences one per row using delimiters (?, ., !)
x <- x %>% dplyr::mutate(!!sym(text_field) := stringr::str_split(!!sym(text_field), "[.!?]", simplify = FALSE)) %>% unnest(!!sym(text_field))

#Omit empty rows, Not NA & Not an Empty String
x <- x %>% dplyr::filter(!is.na(!!sym(text_field)) & trimws(!!sym(text_field)) != "")

#Expand contractions
#Include function as internal 'repolacements_25` in package


#Add `sent_num` sequence variable looping over each document (starts new at switch, break on factor level change for doc_id)
x <- x %>% dplyr::group_by(!!sym(doc_id_field)) %>% 
  dplyr::mutate(Sent_Num = row_number()) %>% dplyr::ungroup()




#Omit all non-alphabetic characters EXCEPT for whitespace
x <- x %>% mutate(!!sym(text_field) := str_replace_all(!!sym(text_field), "[[:punct:]]", "")) 

#Count number of documents (levels of factor doc_id_field)
doc_count <- nlevels(x[[doc_id_field]])

#Sentence count per document (rows per document)

#Generate count of words per sentence


#Write results to empty dataframe which we will continue to populate dataframe_1
summary_table <- data.frame(Measure = character(), MeanTotal = numeric(), SD= numeric(), Min= numeric(), Max=numeric()) 
  
#Split and unlist data to one word-per row format and write to dataframe_2


# Generate Word_Num variable looping down dataframe, break and restart on new factor level of doc_id
 
 #bind dataframe_1 and dataframe_2 into a list, return the list of two dataframes

  return(x)
}
```


# Test clean_1
```{r}
try_it_out <- clean_1(Test_Text1, text_field='my_text', doc_id_field='doc_id')
print(try_it_out)
write.csv(try_it_out, "output/try_it_out.csv")
```



# RESOLVED
1. Add informative error messages explaining to user `doc_id` and `my_text` fields <br>
2. Generate `sentence_num` using seq_along in the first steps after tokenizing by sentence but before splitting into one word per row format <br>
3. Generate `word_num` using seq_along after splitting and unlisting. Make sure to generate this on the grouped data (word_num sequentially ordered by document) <br>
4. Count number of documents in corpus using levels, write to empty dataframe <br>
